Effects of bariatric surgery on functional connectivity of the reward and default mode network: a pre-registered analysis

### 2021-03-22 manuscript/draft_rsfmri.Rmd

Obesity imposes serious health risks and involves alterations in resting-state functional connectivity  of brain networks involved in eating behavior. Bariatric surgery is an effective treatment, but its effects on functional connectivity are still under debate. In this pre-registered study, we investigated the effects of bariatric surgery on major resting-state brain networks (reward and default mode network) in a longitudinal controlled design. 
33 bariatric surgery patients and 15 obese waiting-list control patients (37 females; aged 44.15 ± 11.86 SD years (range 21-68)) underwent magnetic resonance imaging at baseline, after 6 and 12 months. We conducted a pre-registered whole-brain time-by-group interaction analysis, and a time-by-group interaction analysis on within-network connectivity (https://osf.io/f8tpn/, https://osf.io/59bh7/). In exploratory analyses, we investigated the effects of weight loss and head motion. 
Bariatric surgery compared to waiting did not significantly affect functional connectivity (FWE-corrected $p > .05$), neither whole-brain nor within-network. In exploratory analyses, surgery-related BMI decrease (FWE-corrected $p = .041$) and higher average head motion (FWE-corrected $p = .021$) resulted in significantly stronger connectivity of the reward network with medial posterior frontal regions. 
This pre-registered well-controlled study did not support a strong effect of bariatric surgery, compared to waiting, on major resting-state brain networks after 6 months. Exploratory analyses indicated that head motion might have confounded the effects. Data pooling and more rigorous control of within-scanner head motion during data acquisition are needed to substantiate effects of bariatric surgery on brain organization.

@@

The study protocol was in accordance with the Helsinki Declaration and approved by the ethics committee of the Charité University Medicine Berlin. The study was registered at clinicaltrials.gov as NCT01554228. We pre-registered the present resting-state analyses on the OSF (https://osf.io/yp42s). We made additional changes (see https://osf.io/59bh7/) to the pre-registration after preprocessing the rsfMRI data as we realized some aspects of the analysis were inadequately described in the initial pre-registration. For a comparison of the pre-registration and the manuscript, please visit https://osf.io/45n9f/.\
Participants were recruited from the Center for Bariatric and Metabolic Surgery at the Charité University Medicine Berlin. Inclusion criteria were, in accordance with German guidelines for bariatric surgery, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g. type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013]. Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (defined as MMSE score < $34$).\

@@

`r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="RYGB" & !is.na(subj_df$interv),])` patients underwent RYGB, `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="VSG" & !is.na(subj_df$interv),])` underwent VSG and `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="GB" & !is.na(subj_df$interv),])` GB, for `r nrow(subj_df[subj_df$condition=="IG" & is.na(subj_df$interv),])` patients in the intervention group information was not available. \

@@

MRI was performed with a 12-channel head coil on a 3 Tesla Trio, Siemens (Erlangen) with syngo B17 software.
<!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, repetition time (TR) = 1900 ms, echo time (TE) = 2.52 ms, flip angle = 9°, voxel size = 1 x 1 x 1 mm$^3$, 192 sagital slices).
<!-- rsfMRI --> Resting-state echo-planar imaging  was acquired with a TR of 2.3 s and TE of 30 ms. The image matrix was 64 x 64 with an in-plane resolution of 3 mm x 3 mm and 34 slices with a slice thickness of 4 mm. 150 volumes were acquired, resulting in a total acquisition time of 5:45 minutes. Additionally, a gradient echo field map with a TE difference of 2.46 ms was acquired to correct for field inhomogeneities.
Participants were instructed to close their eyes but to remain awake during scanning.

@@

### Minimal preprocessing
Imaging data analysis was conducted using AFNI 19.1.05, ANTS 2.3.1, FSL 6.0.1 and FreeSurfer 6.0.0p1, wrapped in a nipype workflow (version 1.2.0) in Python 2.7.15 which can be found on https://github.com/fBeyer89/ADI_preproc/.
T1-weighted images were first processed by FreeSurfer's cross-sectional pipeline [@fs_cross]. Then, Freesurfer's longitudinal stream was applied to all cross-sectional runs [@fs_long]. Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's `aseg.mgz` segmentation file for quality control of rsfMRI preprocessing.  
The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 1x1x1x mm template using ANTS [@avants2009ants]. 
Minimal functional preprocessing included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`). In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step. 
For further analysis and ICA-AROMA processing, the minimally preprocessed data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).

@@

## Functional connectivity (FC)
To derive RN and DMN FC maps, we used NAcc and PCC/precuneus as seed regions of interest (ROI), respectively. We did not select vmPFC for the RN due to low SNR. Based on FreeSurfer's aseg.mgz and Desikan-Killiany parcellation, we created seed masks using mri_binarize (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres. Then, we used NiftiLabelsMasker and NiftiMasker to extract the standardized timeseries from the seed regions and the whole brain. We calculated the Pearson’s correlation between them with numpy.dot, performed r-to-z Fisher-transformation and saved the resulting correlation maps for each preprocessing pipeline (minimally preprocessed,AROMA, AROMA+CC, AROMA+CC+GSR). Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.

@@

Statistical analyses were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, Ashburner et al. -@spm12). The marginal model implemented in the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term. We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size. We used an explicit brain mask, derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to 3 x 3 x 3 mm$^3$ and thresholded at 0.5 GM probability) for all analyses. Statistical analyses on the aggregated FC and imputation of missing data were performed in R version 3.6.1 [@team2013r].

<!-- MODEL SETUP --> 
### Confirmatory analyses
**Whole-brain analysis**
We tested the pre-registered hypothesis of a time-by-group interaction:  `FC=group + time + group * time`.\
As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (model CA1), age, sex and average logmFD of both timepoints (model CA2) and age, sex, average logmFD and baseline BMI (model CA3). 
Because the information about the BMI at baseline of one participant in the intervention group was missing, we employed multivariate imputation using the package `mice` based on subject identification, condition, baseline age, and available BMI values. This package allows taking into account the multi-level structure of the data and averaging across multiple possible values, in this case 50 different values, each derived from 10 iterations.\
**Aggregated FC analysis**
We analyzed the aggregate FC using linear mixed models in the R package lme4 [@Bates_2007]. We deviated from the pre-registration by only investigating data from the AROMA+CC+GSR denoising pipeline. First, we investigated the time-by-group interaction for baseline and followup time points only. We either performed no correction (model CA1), adjusted for baseline age, sex, average of logmFD of both time points (model CA2) or additionally adjusted for baseline BMI (model CA3). We performed model comparison between R1 and R0 models, where $R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj))$, and $R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj))$. As specified in the pre-registration, we repeated the above-mentioned interaction analysis for all three time points.

@@

**Whole-brain analysis**
<!-- INFERENCE --> To ensure robustness of our results, we used non-parametric inference testing based on wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference. Deviating from the pre-registration, we used type C2 instead of type 3 for small sample bias adjustment as this was recommended for wild bootstraps in the SwE manual. Opposed to the description in the pre-registration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01$, and 1000 bootstraps due to required computation time (instead of 5000)). Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$. The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.\
**Aggregated FC analysis**
The interaction effect of group and time point in the models of aggregated FC was considered significant if the model comparison between R1 and R0 models using the anova command showed $p < 0.05$. In all exploratory models, we considered all coefficients with $p < 0.05$ as significant.

@@

## Exploratory Analyses (pre-registered)
### Whole-brain analysis
RN FC was not significantly related to neither average nor change in BMI, for either of the denoising pipelines and regardless whether we adjusted for logmFD in models EA1 and EA2.
Only in model EA3 (adjusting for average and change in both BMI and logmFD), we found that more BMI decrease (e.g. weight loss) predicted higher FC between NAcc and a cluster in the posterior-medial frontal region (see Figure \@ref(fig:BMIFDagesexNacccgnBMIdeact) and Table \@ref(tab:tableBMIFDmodel)).\
The peak voxel was classified as belonging to superior frontal gyrus ($45$% probability) and supplementary motor area (SMA) ($37$%  probability) in the Harvard-Oxford atlas. Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.030$), also in AROMA+CC+GSR-denoised data ($p = 0.041$).\
```{r "BMIFDagesexNacccgnBMIdeact", fig.cap="Stronger BMI decrease is associated with increased FC between NAcc and posterior-medial frontal region, adjusted for age, sex, average BMI and logmFD (Model EA3)", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
                          "../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact.pdf"))
```
```{r "BMIFDagesexNaccavgFDact", fig.cap="Higher average logmFD is positively associated with FC between NAcc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD (Model EA3)", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
                          "../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf"))
```
Moreover, average logmFD was positively associated with FC between NAcc and motor cortex in model EA3 for both denoising pipelines (see Figure \@ref(fig:BMIFDagesexNaccavgFDact) and Table \@ref(tab:tableBMIFDmodel)).\
For the DMN, we found that higher average BMI predicted lower FC of the precuneus/PCC with the lingual gyrus, mid orbital gyrus and temporal gyrus in the images denoised with AROMA+CC. This finding was significant in models EA1, EA2 and EA3 (see Figures \@ref(fig:BMIagesexPCCccavgBMIdeact) and \@ref(fig:BMIFDagesexPCCccavgBMIdeactt) and Table \@ref(tab:tableBMImodel)). Yet, none of the clusters survived when using AROMA+CC+GSR denoised data (see Table \@ref(tab:tableBMIFDmodel).\
```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Higher average BMI is associated with lower FC of PCC/precuneus with different regions in AROMA+CC denoised data", fig.subcap = c("adjusted for age and sex (Model EA1)", "adjusted for age, sex and logmFD (EA2)"), fig.hold=TRUE, out.width = "50%", fig.align="center"}
knitr::include_graphics(c("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf", "../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
```
```{r "BMIFDagesexPCCccavgBMIdeactt", fig.cap="Higher average BMI is associated with lower FC of  PCC/precuneus with different regions, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data (Model EA3)", out.width = "50%", fig.align="center"}
knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
```
<!-- Neurovault -->
Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average logmFD and change in logmFD are published on NeuroVault [https://neurovault.org/collections/9426/](https://neurovault.org/collections/9426/).

@@

### Average and change in logmFD 
<!-- #### Reward -->As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNACCccavgFDact), higher average logmFD was associated with stronger FC between the NAcc and a cluster located in proximity to the central sulcus and motor areas. This Cluster only differed in size between denoising pipelines.\
We did not find any clusters with a significant association of logmFD and DMN FC.
```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of NAcc with a cluster in motor cortex, adjusted for age, sex, average BMI, change BMI and change logmFD", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
knitr::include_graphics("../report/fig/FDagesex_NACCcc_avgFD_act.pdf")
knitr::include_graphics("../report/fig/FDagesex_NACCgsr_avgFD_act.pdf")
```
### Functional decoding
Functional decoding of the two activation patterns from the pre-registered, exploratory analysis EA3 showed different top association terms. For the [https://neurosynth.org/decode/?neurovault=441783](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula and inferior frontal were the top terms, while for the [https://neurosynth.org/decode/?neurovault=441784](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and premotor cortex were the meta-analytic activation maps most similar to this contract. The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.

>>>>

Obesity imposes serious health risks and involves alterations in resting-state functional connectivity  of brain networks involved in eating behavior. Bariatric surgery is an effective treatment, but its effects on functional connectivity are still under debate. In this pre-registered study, we investigated the effects of bariatric surgery on major resting-state brain networks (reward and default mode network) in a longitudinal controlled design. 
33 bariatric surgery patients and 15 obese waiting-list control patients (37 females; aged 44.15 ± 11.86 SD years (range 21-68)) underwent magnetic resonance imaging at baseline, after 6 and 12 months. We conducted a pre-registered whole-brain time-by-group interaction analysis, and a time-by-group interaction analysis on within-network connectivity (https://osf.io/f8tpn/, https://osf.io/59bh7/). In exploratory analyses, we investigated the effects of weight loss and head motion. 
Bariatric surgery compared to waiting did not significantly affect functional connectivity (FWE-corrected $p > 0.05$), neither whole-brain nor within-network. In exploratory analyses, surgery-related BMI decrease (FWE-corrected $p = 0.041$) and higher average head motion (FWE-corrected $p = 0.021$) resulted in significantly stronger connectivity of the reward network with medial posterior frontal regions. 
This pre-registered well-controlled study did not support a strong effect of bariatric surgery, compared to waiting, on major resting-state brain networks after 6 months. Exploratory analyses indicated that head motion might have confounded the effects. Data pooling and more rigorous control of within-scanner head motion during data acquisition are needed to substantiate effects of bariatric surgery on brain organization.

@@

The study protocol was in accordance with the Helsinki Declaration and approved by the ethics committee of the Charité University Medicine Berlin. The study was registered at clinicaltrials.gov as NCT01554228. We pre-registered the present resting-state analyses on the OSF (https://osf.io/yp42s). We made additional changes (see https://osf.io/59bh7/) to the pre-registration after preprocessing the rsfMRI data as we realized some aspects of the analysis were inadequately described in the initial pre-registration. For a comparison of the pre-registration and the manuscript, please visit https://osf.io/45n9f/. \
Participants were recruited from the Center for Bariatric and Metabolic Surgery at the Charité University Medicine Berlin. Inclusion criteria were, in accordance with German guidelines for bariatric surgery, a failure of conservative obesity treatment and either (1) a BMI $> 40$ km/m$^2$ or (2) a BMI $> 35$ kg/m$^2$ and at least one typical co-morbidity (e.g. type-2 diabetes, hypertension, non-alcoholic fatty liver disease)[@Mechanick_2013]. Participants were aged between $18$ and $70$ years and had no history of cancer, chronic inflammatory disease and addiction, other severe untreated diseases, brain pathologies identified in the MRI scan or cognitive impairments (defined as MMSE score $< 34$).\

@@

`r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="RYGB" & !is.na(subj_df$interv),])` patients underwent RYGB, `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="VSG" & !is.na(subj_df$interv),])` underwent VSG and `r nrow(subj_df[subj_df$condition=="IG" & subj_df$interv=="GB" & !is.na(subj_df$interv),])` GB, for `r nrow(subj_df[subj_df$condition=="IG" & is.na(subj_df$interv),])` patients in the intervention group information was not available. 

@@

MRI was performed with a 12-channel head coil on a 3 Tesla Trio, Siemens (Erlangen) with syngo B17 software.
<!-- MRI --> T1-weighted anatomical images were acquired as described in @Prehn_2020 (with MPRAGE, repetition time (TR) = 1900 ms, echo time (TE) = 2.52 ms, flip angle = 9°, voxel size = $1 \times 1 \times 1$ mm$^3$, 192 sagital slices).
<!-- rsfMRI --> Resting-state echo-planar imaging  was acquired with a TR of 2.3 s and TE of 30 ms. The image matrix was $64 \times 64$ with an in-plane resolution of 3 mm $\times$ 3 mm and 34 slices with a slice thickness of 4 mm. 150 volumes were acquired, resulting in a total acquisition time of 5:45 minutes. Additionally, a gradient echo field map with a TE difference of 2.46 ms was acquired to correct for field inhomogeneities.
Participants were instructed to close their eyes but to remain awake during scanning.

@@

### Minimal preprocessing
Imaging data analysis was conducted using AFNI 19.1.05, ANTS 2.3.1, FSL 6.0.1 and FreeSurfer 6.0.0p1, wrapped in a nipype workflow (version 1.2.0) in Python 2.7.15 which can be found on https://github.com/fBeyer89/ADI_preproc/.
T1-weighted images were first processed by FreeSurfer's cross-sectional pipeline [@fs_cross]. Then, Freesurfer's longitudinal stream was applied to all cross-sectional runs [@fs_long]. Here, white matter (WM) and cerebral spinal fluid (CSF) masks were derived based on FreeSurfer's `aseg.mgz` segmentation file for quality control of rsfMRI preprocessing.  
The skull-stripped brain (`brainmask.mgz`) was then coregistered to the MNI152 $1 \times 1 \times 1$ mm template using ANTS [@avants2009ants]. 
Minimal functional preprocessing included the removal of first four volumes, motion correction (FSL's `MCFLIRT`), fieldmap distortion correction (FSL's `fsl_prepare_fieldmap` and `FUGUE`) and coregistration to the subject’s individual longitudinal anatomical space (FreeSurfer's `bbregister`). In more detail, the transformations derived from the latter three steps were combined into one and applied in a single step. 
For further analysis and ICA-AROMA processing, the minimally preprocessed data were intensity normalized and smoothed with a 6mm Gaussian kernel (`fslmaths -kernel gauss 2.548`).

@@

## Functional connectivity (FC)
To derive RN and DMN FC maps, we used NAcc and PCC/precuneus as seed regions of interest (ROI), respectively. We did not select vmPFC for the RN due to low SNR. Based on FreeSurfer's aseg.mgz and Desikan-Killiany parcellation, we created seed masks using `mri_binarize` (thresholded for NAcc at 26,58; precuneus at 1025,2025) and averaged them over hemispheres. Then, we used NiftiLabelsMasker and NiftiMasker to extract the standardized timeseries from the seed regions and the whole brain. We calculated the Pearson’s correlation between them with numpy.dot, performed r-to-z Fisher-transformation and saved the resulting correlation maps for each preprocessing pipeline (minimally preprocessed, AROMA, AROMA+CC, AROMA+CC+GSR). Finally, the connectivity maps were transformed into MNI space using the affine transformation and non-linear warp derived with ANTS during anatomical preprocessing.

@@

Statistical analyses were performed in MATLAB version 9.7.0.1190202 (R2019b, @MATLAB2018) using the SwE toolbox version 2.2.2 [@Guillaume_2014] as implemented in the Statistical Parametric Mapping software (SPM12.7770, Ashburner et al. -@spm12). The marginal model implemented in the SwE toolbox implicitly accounts for random effects without the need to specify them through the error term. We used a modified SwE assuming different covariance structures for the intervention and the control group because of their unbalanced sample size. We used an explicit brain mask, derived from the MNI ICBM “152 nonlinear 6th generation” atlas (re-sampled to $3 \times 3 \times 3$ mm$^3$ and thresholded at 0.5 GM probability) for all analyses. Statistical analyses on the aggregated FC and imputation of missing data were performed in R version 3.6.1 [@team2013r].

<!-- MODEL SETUP --> 
### Confirmatory analyses
**Whole-brain analysis**
We tested the pre-registered hypothesis of a time-by-group interaction:  `FC = group + time + group * time`.\
As pre-registered, we performed the analysis for baseline and followup time points only, for AROMA+CC and AROMA+CC+GSR denoising pipelines and adjusting for no confounders (model CA1), age, sex and average logmFD of both timepoints (model CA2) and age, sex, average logmFD and baseline BMI (model CA3). 
Because the information about the BMI at baseline of one participant in the intervention group was missing, we employed multivariate imputation using the package `mice` based on subject identification, condition, baseline age, and available BMI values. This package allows taking into account the multi-level structure of the data and averaging across multiple possible values, in this case 50 different values, each derived from 10 iterations.\
**Aggregated FC analysis**
We analyzed the aggregate FC using linear mixed models in the R package lme4 [@Bates_2007]. We deviated from the pre-registration by only investigating data from the AROMA+CC+GSR denoising pipeline. First, we investigated the time-by-group interaction for baseline and followup time points only. We either performed no correction (model CA1), adjusted for baseline age, sex, average of logmFD of both time points (model CA2) or additionally adjusted for baseline BMI (model CA3). We performed model comparison between R1 and R0 models, where `R1 = lmer(aggFC ~ timepoint*group + age + sex + (1|subj))`, and `R0 = lmer(aggFC ~ timepoint + group +  age + sex + (1|subj))`. As specified in the pre-registration, we repeated the above-mentioned interaction analysis for all three time points.

@@

**Whole-brain analysis**
<!-- INFERENCE --> To ensure robustness of our results, we used non-parametric inference testing based on wild bootstrap with an unrestricted SwE on all contrasts of interest for clusterwise inference. Deviating from the pre-registration, we used type C2 instead of type 3 for small sample bias adjustment as this was recommended for wild bootstraps in the SwE manual. Opposed to the description in the pre-registration, we used a cluster forming threshold of $p < 0.001$ for a more rigorous multiple comparison adjustment (instead of $p < 0.01$), and 1000 bootstraps due to required computation time (instead of 5000). Significant clusters are defined as family-wise error (FWE) corrected $p < 0.05$. The anatomical localization of significant clusters was investigated with the SPM Anatomy toolbox, version 2.2c [@Eickhoff_2005] and the Harvard-Oxford Atlas in FSL version 5.0.11.\
**Aggregated FC analysis**
The interaction effect of group and time point in the models of aggregated FC was considered significant if the model comparison between R1 and R0 models using the anova command showed $p < 0.05$. In all exploratory models, we considered all coefficients with $p < 0.05$ as significant.

@@

## Exploratory Analyses (pre-registered)
### Whole-brain analysis
RN FC was not significantly related to neither average nor change in BMI, for either of the denoising pipelines and regardless whether we adjusted for logmFD in models EA1 and EA2.
Only in model EA3 (adjusting for average and change in both BMI and logmFD), we found that more BMI decrease (e.g. weight loss) predicted higher FC between NAcc and a cluster in the posterior-medial frontal region (see Figure \@ref(fig:BMIFDagesexNacccgnBMIdeact) and Table \@ref(tab:tableBMIFDmodel)).\
The peak voxel was classified as belonging to superior frontal gyrus ($45$% probability) and supplementary motor area (SMA) ($37$%  probability) in the Harvard-Oxford atlas. Voxel activation at local maximum within this cluster was significant at peak level after FWE-correction ($p = 0.030$), also in AROMA+CC+GSR-denoised data ($p = 0.041$).\
```{r "BMIFDagesexNacccgnBMIdeact", fig.cap="Stronger BMI decrease is associated with increased FC between NAcc and posterior-medial frontal region, adjusted for age, sex, average BMI and logmFD (Model EA3). Legend denotes empirical Z-values.", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", fig.hold=TRUE, eval=TRUE}
knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_cngBMI_deact.pdf",
                          "../report/fig/BMIFDagesex_NACCgsr_cngBMI_deact.pdf"))
```
```{r "BMIFDagesexNaccavgFDact", fig.cap="Higher average logmFD is positively associated with FC between NAcc and motor cortex, adjusted for age, sex, average BMI, change in BMI, and change in logmFD (Model EA3). Legend denotes empirical Z-values.", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data"), out.width = "50%", eval=TRUE}
knitr::include_graphics(c("../report/fig/BMIFDagesex_NACCcc_avgFD_act.pdf",
                          "../report/fig/BMIFDagesex_NACCgsr_avgFD_act.pdf"))
```
Moreover, average logmFD was positively associated with FC between NAcc and motor cortex in model EA3 for both denoising pipelines (see Figure \@ref(fig:BMIFDagesexNaccavgFDact) and Table \@ref(tab:tableBMIFDmodel)).\
For the DMN, we found that higher average BMI predicted lower FC of the precuneus/PCC with the lingual gyrus, mid orbital gyrus and temporal gyrus in the images denoised with AROMA+CC. This finding was significant in models EA1, EA2 and EA3 (see Figures \@ref(fig:BMIagesexPCCccavgBMIdeact) and \@ref(fig:BMIFDagesexPCCccavgBMIdeactt) and Table \@ref(tab:tableBMImodel)). Yet, none of the clusters survived when using AROMA+CC+GSR denoised data (see Table \@ref(tab:tableBMIFDmodel).\
```{r "BMIagesexPCCccavgBMIdeact",fig.cap="Higher average BMI is associated with lower FC of PCC/precuneus with different regions in AROMA+CC denoised data", fig.subcap = c("adjusted for age and sex (Model EA1). Legend denotes empirical Z-values.", "adjusted for age, sex and logmFD (EA2)"), fig.hold=TRUE, out.width = "50%", fig.align="center"}
knitr::include_graphics(c("../report/fig/BMIagesex_PCCcc_avgBMI_deact.pdf", "../report/fig/BMIagesexfd_PCCcc_avgBMI_deact.pdf"))
```
```{r "BMIFDagesexPCCccavgBMIdeactt", fig.cap="Higher average BMI is associated with lower FC of  PCC/precuneus with different regions, adjusted for age, sex, average BMI, average logmFD, and change in logmFD in AROMA+CC denoised data (Model EA3). Legend denotes empirical Z-values.", out.width = "50%", fig.align="center"}
knitr::include_graphics("../report/fig/BMIFDagesex_PCCcc_avgBMI_deact.pdf")
```
<!-- Neurovault -->
Unthresholded maps for the t-tests as well as contrasts of average BMI and change BMI of the main model, and post-hoc contrasts for average logmFD and change in logmFD are published on NeuroVault [https://neurovault.org/collections/9426/](https://neurovault.org/collections/9426/).

@@

### Average and change in logmFD 
<!-- #### Reward -->As shown in Table \@ref(tab:tableFDmodel) and Figure \@ref(fig:FDagesexNACCccavgFDact), higher average logmFD was associated with stronger FC between the NAcc and a cluster located in proximity to the central sulcus and motor areas. This Cluster only differed in size between denoising pipelines.\
We did not find any clusters with a significant association of logmFD and DMN FC.
```{r "FDagesexNACCccavgFDact", fig.show = "hold", out.width = "50%", fig.cap = "Positive association of average logmFD and FC of NAcc with a cluster in motor cortex, adjusted for age, sex, average BMI, change BMI and change logmFD. Legend denotes empirical Z-values.", fig.subcap = c("in AROMA+CC", "in AROMA+CC+GSR data")}
knitr::include_graphics("../report/fig/FDagesex_NACCcc_avgFD_act.pdf")
knitr::include_graphics("../report/fig/FDagesex_NACCgsr_avgFD_act.pdf")
```
### Functional decoding
Functional decoding of the two activation patterns from the pre-registered, exploratory analysis EA3 showed different top association terms. For the [https://neurosynth.org/decode/?neurovault=441783](https://neurosynth.org/decode/?neurovault=441783) frontal, anterior insula and inferior frontal were the top terms, while for the [https://neurosynth.org/decode/?neurovault=441784](https://neurosynth.org/decode/?neurovault=441784) primary motor, motor and premotor cortex were the meta-analytic activation maps most similar to this contract. The FC contrasts were thus somewhat distinct, though the decoding method [does not allow](https://www.talyarkoni.org/blog/tag/neurosynth/) to conclude specificity.
